---
title: "Writing Reports with R Markdown"
author: "Sarah Katz"
date: "2023-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
```

This report was prepared to the attention of the United Nations. It analyzes the relationship between a country's GDP, its life expectancy and CO2 emission. Our goal is to determine to what degree a county's economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesize that both life expectancy and CO2 emissions increase with s country's GDP.


```{r data, include=FALSE}
gapminder_1997 <- read_csv("data/gapminder_1997.csv")
```
```{r gdp_lifeexp_1997, echo=FALSE}
ggplot(data = gapminder_1997) +
  aes(x = gdpPercap, y = lifeExp, color = continent, size = pop/1000000) +
  geom_point() +
  labs(x = "GDP Per Capita", 
       y = "Life Expectancy", 
       title = "Do people in wealthy countries live longer?", 
       size = "Population (in Millions)")
```


#Use coded variable in line of the text

#defined variable to be used in writing
```{r include=FALSE}
nCountries <- gapminder_1997 %>% 
  select(country) %>% 
  n_distinct()

minGDP <- gapminder_1997 %>% 
  summarise(round(min(gdpPercap))) %>% 
  pull()

maxGDP <- gapminder_1997 %>% 
  summarise(round(max(gdpPercap))) %>% 
  pull()

```


The above plot shows the relationship between GDP per capita and life expectancy for a total of `r nCountries ` countries. For this set of countries, economic wealth ranged from a minimum of USD `r minGDP ` to a maximum of USD `r maxGDP `

#Creating a table

|HEADER 1|HEADER 2|
|--------|--------|
|row 1, column 1|row 1, column 2|
|row 2, column 1|row 2, column 2|


|Summary of Data|
|--------|--------|
|Number of Countries|`r nCountries`|
|Minimum GDP Per Capita|`r minGDP`|
|Maximum GDP Per Capita|`r maxGDP`|


```{r kabletable}
library(knitr)

gapminder_1997 %>% 
  filter(continent == "Oceania") %>% 
  kable()
```

# Header 1
## Header 2
### My third header

Exercise: R Markdown Synax
* Create a bullet point list with three items
* as the first item, write the name of your currently favorite programing language in bold
* as the second item, write the name of a function you have so far found most useful
* as the third item, write one thing you want to learn on your programming journey in bold italics
* turn your bullet point list into a numbered list
* create a forth list item and find an online guide and/or cheat sheet for basic Markdown synax, write its name down here and hyperlink its url

My List:
* **R**
* *ggplot()*
* ***I want to know more about plotting data***

1. **R**
2. *ggplot()*
3. ***I want to know more about plotting data***
4. [https://www.markdownguide.org/cheat-sheet/](Great Markdown Guide)
5. [https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf]








